---
title: "Characteristics of Different Democracy Measures"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Characteristics of Different Democracy Measures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
library(knitr)
library(QuickUDS)
library(dplyr)
library(ggplot2)
library(GGally)
opts_chunk$set(echo=TRUE)
options(knitr.table.format = 'markdown')

```

This vignette describes the temporal and spatial coverage of the democracy measures included in this package,  notes their correlations, and documents any changes made to the original data sources.

# General Characteristics

```{r general_characteristics}
library(knitr)
library(QuickUDS)
library(dplyr)
library(reshape2)

democracy_long <- democracy %>%
        melt(id.vars = 1:15, na.rm=TRUE) 

kable(democracy_long %>%
        group_by(variable) %>%
        summarise(distinct_countries = n_distinct(country_name),
                  distinct_years = n_distinct(year),
                  min_year = min(year),
                  max_year = max(year),
                  mean_year = mean(year),
                  num_values = n_distinct(value),
                  mean = mean(value),
                  median = median(value),
                  min_value = min(value),
                  max_value = max(value),
                  sd = sd(value)),
      digits = 2)

```


# Arat

This is the dataset described in Arat 1991; the actual values are taken from Pemstein, Meserve, and Melton 2013 (the replication data for Pemstein, Meserve, and Melton 2010).

## Coverage

```{r arat_coverage, fig.width=7,fig.height=7}

library(ggplot2)

panel <- democracy %>% select(country_name,year) %>% distinct()

temporal_coverage <- function(data) {
  data <- left_join(panel,data) 
  
  data <- data %>% 
    group_by(year, add=TRUE) %>% 
    count(year,in_system)

  ggplot(data =  data, aes(x=year,fill = in_system, y = n)) + 
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(fill = "In G&W system of states", 
         y = "Number of states or territories included\nin each year of `democracy` panel") +
    ggtitle("Coverage \nwithin the `democracy` dataset")
}

library(rworldmap)

world <- getMap()
world <- fortify(world)

spatial_coverage <- function(data) {
  ggplot() +
    geom_path(data = world, aes(x=long,y=lat,group=group)) +
    theme_bw() +
    theme(legend.position = "bottom") +
    geom_count(data = data, aes(x=lon,y=lat,color = in_system)) +
    labs(color = "In G&W system of states", y = "", x = "", size = "Number of country-years") +
    ggtitle("Spatial coverage")
}

data <- democracy_long %>% 
  filter(variable == "arat_pmm")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r arat_distribution}
ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)

```

# BLM

This is the dataset described in Bowman, Lehoucq, and Mahoney 2005.

## Coverage

```{r blm_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "blm")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r blm_distribution}
ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# BMR

This is the dataset described in Boix, Miller, and Rosato 2012.

## Coverage

```{r bmr_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "bmr_democracy")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r bmr_distribution}
data <- democracy_long %>% filter(variable %in% c("bmr_democracy","bmr_democracy_omitteddata"))

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# BNR

This is the dataset described in Bernhardt, Nordstrom, and Reenock 2001. The original dataset only counts periods of democracy in the period 1913-2005, since it is designed for event history analysis. To put it in country-year format, this package assumes that country-years in independent states between 1913 and 2005 are to be counted as non-democratic if they are not explicitly said to be democratic by BNR. (Country-years are considered to be independent if the state is not a microstate and appears in Gleditasch and Ward's [1999] panel of indepdent states for the period).

## Coverage

```{r bnr_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "bnr")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r bnr_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Bollen

This is the dataset described in Bollen 2001. The actual values are taken from Pemstein, Meserve, and Melton's replication data for their article (Pemstein, Meserve, and Melton 2013).

## Coverage

```{r bollen_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "bollen_pmm")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r bollen_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)

```

# Doorenspleet

This is the dataset described in Doorenspleet 2000.

## Coverage

```{r doorenspleet_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "doorenspleet")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r doorenspleet_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# EIU

This contains the Economist Intelligence Unit's index of democracy (EIU 2012). The actual data is taken from the World Bank's Governance Indicators (http://www.govindicators.org)

## Coverage

```{r eiu_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "eiu")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r eiu_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)
```

# Freedom House (Freedom in the World)

This is the Freedom in the World Index, to 2015 (Freedom House 2016). Some non-independent territories have been excluded from the original data.

## Coverage

```{r fh_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable %in% c("freedomhouse"))

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r fh_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)

```

# Freedom House (Electoral Democracies)

This is Freedom House's list of electoral democracies, to 2015 (Freedom House 2016). Some non-independent territories have been excluded from the original data.

## Coverage

```{r fh_electoral_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "freedomhouse_electoral")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r fh_electoral_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# GWF (Geddes, Wright, and Frantz)

This is a measure of democracy/non-democracy derived from the dataset described in Geddes, Wright, and Frantz 2014. The original data has been extended beyond 1945 by reconciling the information contained in the original dataset's `gwf_startdate`, `gwf_spell`, and `gwf_casename` variables, which encode the start year of each democratic and non-democratic regime (sometimes going back to the 18th century). 

## Coverage

```{r gwf_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "gwf")

temporal_coverage(data) +
  geom_vline(xintercept = 1945) +
  annotate("text", label = "Limit of original dataset", x = 1945,y = 100, angle=90)

spatial_coverage(data)
```

## Distribution

```{r gwf_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```


# Hadenius

This is the dataset in Hadenius 1992. Actual values taken from Pemstein, Meserve, and Melton 2013. 

## Coverage

```{r hadenius_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "hadenius_pmm")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r hadenius_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value),bins=20) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)
  
```

# Kailitz

The dataset described in Kailitz 2013.  

## Coverage

```{r kailitz_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "kailitz_binary")

temporal_coverage(data)

spatial_coverage(data)

```

## Distribution

```{r kailitz_distribution }

data <- democracy_long %>% filter(variable %in% c("kailitz_binary","kailitz_tri"))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

## Index construction and known problems

Note that 316 of the country-years in the Kailitz dataset are classified with more than one regime type. 

```{r kailitz_classification_problems_1}
kable(kailitz_yearly %>% 
        count(multiple_regimes = grepl("-",combined_regime)))

kable(kailitz_yearly %>% 
        filter(grepl("-",combined_regime)) %>%
        group_by(country_name) %>%
        arrange(country_name,year) %>%
        group_by(combined_regime, add=TRUE) %>%
        summarise(min = min(year), max = max(year), n = n()))

```

The following are especially troublesome, since the multiple categories do not make sense:

```{r kailitz_classification_problems_2}
kable(kailitz_yearly %>% 
        filter(grepl("-",combined_regime)) %>%
        group_by(country_name) %>%
        arrange(country_name,year) %>%
        group_by(combined_regime, add=TRUE) %>%
        summarise(min = min(year), max = max(year), n = n()) %>%
        filter(grepl("democracy",combined_regime, ignore.case=TRUE)))
```

I have constructed the index to classify a country as "democratic" only if it is not also classified as a non-democratic regime as well. Here are the index counts for each regime type:

```{r kailitz_index_count}
kable(kailitz_yearly %>%
        count(kailitz_binary,kailitz_tri,combined_regime))
```

# LIED

This is the Lexical Index of Democracy described in Skaaning,  Gerring, and Bartusevicius 2015 (version 3, updated to 2015). 

## Coverage

```{r lied_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "lied")

temporal_coverage(data)

spatial_coverage(data)
```

## Distribution

```{r lied_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value),binwidth=1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Magaloni

This is a measure of democracy derived from the authoritarian regimes dataset in Magaloni, Chu, and Min 2013. The original dataset has been extended beyond 1950 using the information encoded in the `duration_nr` variable of the original dataset, which provides information about the start date of each regime. 

## Coverage

```{r magaloni_extended_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "magaloni_democ_binary")

temporal_coverage(data) +
  geom_vline(xintercept = 1950) +
  annotate("text", label = "Limit of original dataset", x = 1950,y = 100, angle=90)

spatial_coverage(data)

```

## Distribution

One change has been made (Pakistan 1971/1972 appears to have been misclassified as a democracy).

Note `magaloni_regime_tri` identifies as "hybrid" (middle level) all multiparty autocracies.

```{r magaloni_distribution }

data <- democracy_long %>% filter(variable %in% c("magaloni_democ_binary","magaloni_regime_tri"))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)
  
```

# Mainwaring

This is the dataset in Mainwaring, Brinks, and Perez Linan 2008.

## Coverage

```{r mainwaring_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "mainwaring")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r mainwaring_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Munck

This is the dataset in Munck 2009. Taken from Pemstein, Meserve, and Melton 2013.

## Coverage

```{r munck_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "munck_pmm")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r munck_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)
```

# PACL/ACLP/DD

This is the dataset described in Cheibub, Gandhi, and Vreeland 2010.

## Coverage

```{r pacl_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "pacl")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r pacl_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years", x="") +
  facet_wrap(~variable, ncol=2) 

```

# PEPS

This is the dataset described in Moon et al 2006. 

## Coverage

The `democracy` file contains seven variables from this dataset: `PEPS1i`, `PEPS2i`, `PEPS1q`, `PEPS2q`, `PEPS1v`,`PEPS2v`, and `Polity3`, a cleaned up version of the `polity2` variable in the Polity IV data. The `PEPS*` variables are constructed from `Polity3` and a participation variable derived either from Vanhanen's population and raw participation data (variables ending in `*v`) or from voting turnout data from IDEA (variables ending in `q` or `i`).

```{r peps_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable %in% c("PEPS1i","PEPS2i","PEPS1q","PEPS2q","PEPS1v","PEPS2v","Polity3"))

data <- left_join(panel,data) 
  
data <- data %>% group_by(variable,year, add=TRUE) %>% count_(vars = c("year","in_system","variable"))

ggplot(data =  data, aes(x=year,fill = in_system, y = n)) + 
    geom_bar(stat = "identity") +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(fill = "In G&W system of states", y = "Number of states or territories\nin each year of `democracy` panel") +
    ggtitle("Temporal coverage \nwithin the `democracy` dataset") + 
  facet_wrap(~variable) 

data <- democracy_long %>% filter(variable %in% c("PEPS1i","PEPS2i","PEPS1q","PEPS2q","PEPS1v","PEPS2v"))

spatial_coverage(data)
```

## Distribution

```{r peps_distribution, fig.width= 5,fig.height=5}

ggplot(data = data) +
  geom_histogram(aes(x=value),binwidth=1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years", x="") +
  facet_wrap(~variable, ncol=2)

library(GGally)
ggcorr(data = democracy %>% select(PEPS1i:PEPS2v), label=TRUE,label_round=3) + scale_fill_gradient2(midpoint = 0.7)

```

## Known problems

The `Polity3` variable is different from the `polity2` variable in the following cases, mostly due to the different way in which Moon et al recode transitional periods, but in some cases due to revisions in the underlying Polity IV dataset since 2006:

```{r peps_problems}
kable(democracy %>%
        filter(Polity3 != polity2) %>%
        group_by(country_name,Polity3,polity2,polity) %>%
        summarise(years = min(year), max = max(year), n = n()))

```

# PITF

These are the scores used in Goldstone et al 2010 and (in binary form) in Taylor and Ulfelder 2015. 

## Coverage

```{r pitf_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "pitf")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r pitf_distribution }

ggplot(data = data) +
  geom_bar(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years", x="") +
  facet_wrap(~variable, ncol=2) 

data <- democracy_long %>% filter(variable == "pitf_binary")

ggplot(data = data) +
  geom_bar(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years", x="") +
  facet_wrap(~variable, ncol=2) 

```

# Polity

These are variables from the Polity IV dataset in country-year format, updated to 2014 (Marshall, Gurr, and Jaggers 2012). 

## Coverage


```{r polity_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "polity")

temporal_coverage(data) 

spatial_coverage(data)
```

## Distribution

```{r polity_distribution}

data <- democracy_long %>% filter(variable == "polity2")

ggplot(data = data) +
  geom_histogram(aes(x=value),binwidth=1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Polyarchy

The dataset from Coppedge and Reinicke 1991, revised in 2003-2006 with a new "contestation" measure. 

## Coverage

```{r polyarchy_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "polyarchy_reversed")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r polyarchy_distribution}

data <- democracy_long %>% filter(variable %in% c("polyarchy_reversed","polyarchy_contestation"))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

kable(democracy %>% 
        filter(!is.na(polyarchy_reversed)) %>% 
        count(polyarchy_reversed,polyarchy_contestation),
      caption = "Relationship between reversed polyarchy measure and polyarchy contestation measure")

```

# Political Regime Change (PRC)/Gasiorowski dataset

The dataset first described in Gasiorowski 1996, and updated and revised in Reich 2002. 

## Coverage

```{r prc_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "prc")

temporal_coverage(data) 

spatial_coverage(data)
```

## Distribution

```{r prc_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# PIPE

A constructed regime variable from Przeworski 2013 (the Political Institutions and Political Events (PIPE) Data Set).

## Coverage

```{r pipe_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "przeworski")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

It is not clear that this index is correctly constructed, given the confusing documentation in the original dataset. Use with care.

```{r pipe_distribution}

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Svolik

This uses a measure of democracy derived from the authoritarian regime dataset in Svolik 2012. The original data is extended to before 1946 using the information encoded in the `o_startdate` variable of the original dataset, which provides information for the start dates of some authoritarian regimes.

## Coverage

```{r svolik_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "svolik")

temporal_coverage(data) +
  geom_vline(xintercept = 1946) +
  annotate("text", label = "Limit of original dataset", x = 1946,y = 100, angle=90)

spatial_coverage(data)

```

## Distribution

```{r svolik_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)
  
```

# Ulfelder

The dataset in Ulfelder 2012. 

## Coverage

```{r ulfelder_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "ulfelder")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r ulfelder_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)
  
```

# UTIP

A measure of democracy from the regime type dataset described in Hsu 2008. This dataset identifies three types of democracies: "social democracies", "conservative democracies", and "one party democracies." "One party democracies" are poorly documented, but seem to be equivalent to multiparty autocracies. `utip_dichotomous_strict` identifies as democracies only social or conservative democracies; `utip_dichotomous` also identifies as democracies those "one party democracies"; and `utip_trichotomous` assumes that "one party democracies" are hybrid regimes between democracy and non-democracy.

## Coverage

```{r utip_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "utip_dichotomous")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r utip_distribution }

data <- democracy_long %>% filter(variable %in% c("utip_dichotomous_strict","utip_dichotomous","utip_trichotomous"))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)
  
```

# Vanhanen

This is the dataset in Vanhanen 2012. 

## Coverage

```{r vanhanen_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "vanhanen_democratization")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r vanhanen_distribution }

data <- democracy_long %>% filter(variable %in% c("vanhanen_democratization","vanhanen_participation","vanhanen_competition"))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)  
```

# V-Dem

This is a selection of the major democracy indexes from version 6 of the V-Dem dataset (`v2x*` variables) (Coppedge et al. 2016).

## Coverage

Most of the not in-system country-years in the main dataset are from V-Dem.

```{r vdem_coverage, fig.width=7,fig.height=7}
data <- democracy_long %>% filter(variable == "v2x_polyarchy")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r vdem_distribution, fig.width= 5,fig.height=5 }

data <- democracy_long %>% filter(grepl("v2x",variable))

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

ggplot(data = data) +
  geom_density(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Density") +
  facet_wrap(~variable, ncol=2)  

ggcorr(data = democracy %>% select(v2x_api:v2x_polyarchy), label=TRUE,label_round=3, hjust=1) + scale_fill_gradient2(midpoint = 0.7)

```

# Wahman Teorell and Hadenius

This is a measure of democracy from the authoritarian Regimes Data Set, version 5.0, by Hadenius, Teorell, & Wahman, described in Hadenius and Teorell 2007 and in Wahman, Teorell, Hadenius 2013. 

## Coverage

```{r wahman_coverage, fig.width=7,fig.height=7}

data <- democracy_long %>% filter(variable == "wahman_teorell_hadenius")

temporal_coverage(data) 

spatial_coverage(data)

```

## Distribution

```{r wahman_distribution }

ggplot(data = data) +
  geom_histogram(aes(x=value)) +
  theme_bw() +
  theme(legend.position = "bottom") +
  labs(y = "Country-years") +
  facet_wrap(~variable, ncol=2)

```

# Differences between Gleditsch and Ward codes, Polity codes and COW codes

```{r code_differences}
count_sequence_breaks <- function(seq, seq_step = 1) {
  first_diff <- c(seq_step, diff(seq)) - seq_step
  periods <- cumsum(abs(first_diff))
  periods
}

kable(democracy %>%
        filter(GWn != cown | GWn != polity_ccode | cown != polity_ccode) %>%
        distinct(country_name,GWn,cown,polity_ccode,year) %>%
        arrange(country_name,GWn,cown,polity_ccode,year) %>% 
        group_by(country_name,GWn,cown,polity_ccode) %>%
        mutate(period = count_sequence_breaks(year)) %>%
        group_by(period, add = TRUE) %>%
        summarise(min = min(year), max = max(year), n = n()),
      caption = "Differences between Gleditsch and Ward codes, COW codes, and Polity codes")
```

# Differences with the CoW system of states

There are a few country-years that are in the CoW state system but are not included in this dataset:

```{r cow_system_differences}

COW_system <- read.csv("http://www.correlatesofwar.org/data-sets/state-system-membership/system2011/at_download/file")

COW_system <- COW_system %>% rename(cown = ccode)

kable(anti_join(COW_system,democracy))
```

Germany in 1990 appears with the `GWn` code 260 for 1990, since Gleditsch and Ward treats it as a continuation of the Federal Republic; similarly, the Yemen Arab Republic appears with the `GWn` code 678 for 1990, since Gleditsch and Ward treat is a continuation of the previous state, while CoW considers it a new state (code 679). The rest - Zanzibar 1964, Sicily (Kingdom of the Two Sicilies) 1861, Austria-Hungary 1918, Tunisia 1881, and Egypt 1882 have no data for the democracy measures. These cases also represent years where these countries lost their independence or disappeared.

It is also worth noting that, though there are no duplicate country-years wehn grouping by `country_name` or `GWn` or `polity_ccode`, when grouping by `cown`, there are a few duplicate observations:

```{r cow_system_duplicates}
kable(democracy %>% 
        group_by(country_name,year) %>% 
        filter(n() > 1) %>% 
        group_by(country_name) %>% 
        summarise(countries = paste(unique(country_name),collapse = ", "), min(year), max(year)))

kable(democracy %>% 
        group_by(GWn,year) %>% 
        filter(n() > 1) %>% 
        group_by(GWn) %>% 
        summarise(countries = paste(unique(country_name),collapse = ", "), min(year), max(year)))

kable(democracy %>% 
        group_by(polity_ccode,year) %>% 
        filter(n() > 1) %>% 
        group_by(polity_ccode) %>% 
        summarise(countries = paste(unique(country_name),collapse = ", "), min(year), max(year)))

kable(democracy %>% 
        group_by(cown,year) %>% 
        filter(n() > 1) %>% 
        group_by(cown) %>% 
        summarise(countries = paste(unique(country_name),collapse = ", "), min(year), max(year)))
```

# References

Arat, Zehra F. 1991. Democracy and human rights in developing countries.
Boulder: Lynne Rienner Publishers.

Bernhard, Michael Timothy Nordstrom, and Christopher Reenock, "Economic
Performance, Institutional Intermediation and Democratic Breakdown,"
Journal of Politics 63:3 (2001), pp. 775-803. Data and coding description
available at http://users.clas.ufl.edu/bernhard/content/data/data.htm

Boix, Carles, Michael Miller, and Sebastian Rosato. 2012. A Complete Data
Set of Political Regimes, 1800-2007. Comparative Political Studies 46 (12):
1523-1554. Original data available at
https://sites.google.com/site/mkmtwo/democracy-v2.0.dta?attredirects=0


Bollen, Kenneth A. 2001. "Cross-National Indicators of Liberal Democracy,
1950-1990." 2nd ICPSR version. Chapel Hill, NC: University of North
Carolina, 1998. Ann Arbor, MI: Inter-university Consortium for Political
and Social Research, 2001. Original data available at
http://webapp.icpsr.umich.edu/cocoon/ICPSR-STUDY/02532.xml.

Bowman, Kirk, Fabrice Lehoucq, and James Mahoney. 2005. Measuring Political
Democracy: Case Expertise, Data Adequacy, and Central America. Comparative
Political Studies 38 (8): 939-970.
http://cps.sagepub.com/content/38/8/939. Data available at
http://www.blmdemocracy.gatech.edu/.

Cheibub, Jose Antonio, Jennifer Gandhi, and James Raymond Vreeland. 2010.
"Democracy and Dictatorship Revisited." Public Choice. 143(1):67-101.
Original data available at
https://sites.google.com/site/joseantoniocheibub/datasets/democracy-and-dictatorship-revisited.

Coppedge, Michael, John Gerring, Staffan I. Lindberg, Svend-Erik Skaaning,
and Jan Teorell, with David Altman, Michael Bernhard, M. Steven Fish, Adam
Glynn, Allen Hicken, Carl Henrik Knutsen, Kelly McMann, Pamela Paxton,
Daniel Pemstein, Jeffrey Staton, Brigitte Zimmerman, Frida Andersson,
Valeriya Mechkova, Farhad Miri. 2015. V-Dem Codebook v5. Varieties of
Democracy (V-Dem) Project. Original data available at
https://v-dem.net/en/data/.

Coppedge, Michael and Wolfgang H. Reinicke. 1991. Measuring Polyarchy. In
On Measuring Democracy: Its Consequences and Concomitants, ed. Alex
Inkeles. New Brunswuck, NJ: Transaction pp. 47-68.

Doorenspleet, Renske. 2000. Reassessing the Three Waves of Democratization.
World Politics 52 (03): 384-406. DOI: 10.1017/S0043887100016580.
http://dx.doi.org/10.1017/S0043887100016580.

Economist Intelligence Unit. 2012. Democracy Index 2012: Democracy at a
Standstill.

Freedom House. 2015. "Freedom in the World." Original data available at
http://www.freedomhouse.org.

Gasiorowski, Mark J. 1996. "An Overview of the Political Regime Change
Dataset." Comparative Political Studies 29(4):469-483.

Geddes, Barbara, Joseph Wright, and Erica Frantz. 2014. Autocratic
Breakdown and Regime Transitions: A New Data Set. Perspectives on Politics
12 (1): 313-331. Original data available at
http://dictators.la.psu.edu/.

Gleditsch, Kristian S. & Michael D. Ward. 1999. "Interstate System
Membership: A Revised List of the Independent States since 1816."
International Interactions 25: 393-413. The list can be found at
http://privatewww.essex.ac.uk/~ksg/statelist.html

Goldstone, Jack, Robert Bates, David Epstein, Ted Gurr, Michael Lustik,
Monty Marshall, Jay Ulfelder, and Mark Woodward. 2010. A Global Model for
Forecasting Political Instability. American Journal of Political Science 54
(1): 190-208. DOI:10.1111/j.1540-5907.2009.00426.x

Hadenius, Axel. 1992. Democracy and Development. Cambridge: Cambridge
University Press.

Hadenius, Axel & Jan Teorell. 2007. "Pathways from Authoritarianism", Journal
of Democracy 18(1): 143-156.

Hsu, Sara "The Effect of Political Regimes on Inequality, 1963-2002," UTIP
Working Paper No. 53 (2008),
http://utip.gov.utexas.edu/papers/utip_53.pdf. Data available for
download at http://utip.gov.utexas.edu/data/.

Kailitz, Steffen. 2013. Classifying political regimes revisited: legitimation
and durability. Democratization 20 (1): 39-60. Original data available at
http://dx.doi.org/10.1080/13510347.2013.738861.

Mainwaring, Scott, Daniel Brinks, and Anibal Perez Linan. 2008. "Political
Regimes in Latin America, 1900-2007." Original data available from
http://kellogg.nd.edu/scottmainwaring/Political_Regimes.pdf.

Magaloni, Beatriz, Jonathan Chu, and Eric Min. 2013. Autocracies of the
World, 1950-2012 (Version 1.0). Dataset, Stanford University. Original data
and codebook available at
http://cddrl.fsi.stanford.edu/research/autocracies_of_the_world_dataset/

Marshall, Monty G., Ted Robert Gurr, and Keith Jaggers. 2012. "Polity IV:
Political Regime Characteristics and Transitions, 1800-2012." Updated to
2015. Original data available from
http://www.systemicpeace.org/polity/polity4.htm.

Moon, Bruce E., Jennifer Harvey Birdsall, Sylvia Ceisluk, Lauren M.
Garlett, Joshua J. Hermias, Elizabeth Mendenhall, Patrick D. Schmid, and
Wai Hong Wong (2006) "Voting Counts: Participation in the Measurement of
Democracy" Studies in Comparative International Development 42, 2 (Summer,
2006). The complete dataset is available here:
http://www.lehigh.edu/~bm05/democracy/Obtain_data.htm.

Munck, Gerardo L. 2009. Measuring Democracy: A Bridge Between Scholarship
and Politics. Baltimore: Johns Hopkins University Press.

Pemstein, Daniel, Stephen Meserve, and James Melton. 2010. Democratic
Compromise: A Latent Variable Analysis of Ten Measures of Regime Type.
Political Analysis 18 (4): 426-449.

Pemstein, Daniel, Stephen A. Meserve, and James Melton. 2013. "Replication
data for: Democratic Compromise: A Latent Variable Analysis of Ten Measures
of Regime Type." In: Harvard Dataverse.
http://hdl.handle.net/1902.1/PMM

Przeworski, Adam et al. 2013. Political Institutions and Political Events
(PIPE) Data Set. Department of Politics, New York University.
https://sites.google.com/a/nyu.edu/adam-przeworski/home/data

Reich, G. 2002. Categorizing Political Regimes: New Data for Old Problems.
Democratization 9 (4): 1-24.
http://www.tandfonline.com/doi/pdf/10.1080/714000289.

Skaaning, Svend-Erik, John Gerring, and Henrikas Bartusevicius. 2015. A
Lexical Index of Electoral Democracy. Comparative Political Studies 48
(12): 1491-1525. Original data available from
http://thedata.harvard.edu/dvn/dv/skaaning.

Svolik, Milan. 2012. The Politics of Authoritarian Rule. Cambridge and New
York: Cambridge University Press. Original data available from
http://campuspress.yale.edu/svolik/the-politics-of-authoritarian-rule/.

Taylor, Sean J. and Ulfelder, Jay, A Measurement Error Model of Dichotomous
Democracy Status (May 20, 2015). Available at SSRN:
http://ssrn.com/abstract=2726962 or
http://dx.doi.org/10.2139/ssrn.2726962

Ulfelder, Jay. 2012. "Democracy/Autocracy Data Set." In: Harvard Dataverse.
http://hdl.handle.net/1902.1/18836.

Vanhanen, Tatu. 2012. "FSD1289 Measures of Democracy 1810-2012." Original
data available from
http://www.fsd.uta.fi/english/data/catalogue/FSD1289/meF1289e.html

Wahman, Michael, Jan Teorell, and Axel Hadenius. 2013. Authoritarian regime
types revisited: updated data in comparative perspective. Contemporary
Politics 19 (1): 19-34.
